---
title: "2_main_analysis"
author: "Joy_Fu"
date: "2023-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1. Two-sample MR
```{r message=FALSE, warning=FALSE}
rm(list = ls())
lapply(paste('package:', names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE)
pacman::p_load(tidyverse, TwoSampleMR, MRPRESSO, LDlinkR)
# Basic setups
work_data_path = '/Users/Mingzhou/Desktop/AD_Grant/smk_dem_MR/data/'
output_path = '/Users/Mingzhou/Desktop/AD_Grant/smk_dem_MR/output/'
# Source in useful functions
source("functions.R")
```

## 1. Data preparation
bash step - only select significant SNPs from the GSCAN GWASs
awk -F"\t" '{if (NR==1||$8<=5e-08) print $0}' SmokingInitiation.txt > sig_SmokingInitiation.txt (N=7847 SNPs)
awk -F"\t" '{if (NR==1||$8<=5e-08) print $0}' AgeofInitiation.txt > sig_AgeofInitiation.txt (N=450 SNPs)
awk -F"\t" '{if (NR==1||$8<=5e-08) print $0}' SmokingCessation.txt > sig_SmokingCessation.txt (N=224 SNPs)
awk -F"\t" '{if (NR==1||$8<=5e-08) print $0}' CigarettesPerDay.txt > sig_CigarettesPerDay.txt (N=2131 SNPs)

## 2. Perform MR
```{r}
selected_method = c("MR Egger", "Weighted median", "Inverse variance weighted", "Weighted mode")
```

### 1) Smoking initiation
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
exposure_GSCAN_SI = read_exposure_data(paste0(work_data_path, "sumstats/sig_SmokingInitiation.txt"), sep = "\t", 
                                       samplesize_col = "N", snp_col = "RSID", beta_col = "BETA", se_col = "SE", pval_col = "P", 
                                       eaf_col = "AF", effect_allele_col = "REF", other_allele_col = "ALT", clump = TRUE)
dim(exposure_GSCAN_SI) # dim = (93,13)
# Get effects of instruments on outcome
outcome_Kunkle_AD = extract_outcome_data(snps = exposure_GSCAN_SI$SNP, outcomes = "ieu-b-2") # Kunkle,2019
# To detect if there is any IVs not available in outcome data
missing = exposure_GSCAN_SI %>% filter(!SNP %in% outcome_Kunkle_AD$SNP) # 0 SNP missing
# Harmonization of data
dat_SI = harmonise_data(exposure_dat = exposure_GSCAN_SI, outcome_dat = outcome_Kunkle_AD, action = 2)
# Perform MR
mr_SI_result = extract_MR(mr(dat_SI), "smk_init", selected_method)
```

### 2) Cigarettes per day
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
exposure_GSCAN_CPD = read_exposure_data(paste0(work_data_path, "sumstats/sig_CigarettesPerDay.txt"), sep = "\t", 
                                       samplesize_col = "N", snp_col = "RSID", beta_col = "BETA", se_col = "SE", pval_col = "P", 
                                       eaf_col = "AF", effect_allele_col = "REF", other_allele_col = "ALT", clump = TRUE)
dim(exposure_GSCAN_CPD) # dim = (23,13)
# Get effects of instruments on outcome
outcome_Kunkle_AD = extract_outcome_data(snps = exposure_GSCAN_CPD$SNP, outcomes = "ieu-b-2") # Kunkle,2019
# To detect if there is any IVs not available in outcome data
missing = exposure_GSCAN_CPD %>% filter(!SNP %in% outcome_Kunkle_AD$SNP) # 0 SNP missing
# Harmonization of data
dat_CPD = harmonise_data(exposure_dat = exposure_GSCAN_CPD, outcome_dat = outcome_Kunkle_AD, action = 2)
# Perform MR
mr_CPD_result = extract_MR(mr(dat_CPD), "cig_per_day", selected_method)
```

### 3) Age of initiation
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
exposure_GSCAN_AI = read_exposure_data(paste0(work_data_path, "sumstats/sig_AgeofInitiation.txt"), sep = "\t", 
                                       samplesize_col = "N", snp_col = "RSID", beta_col = "BETA", se_col = "SE", pval_col = "P", 
                                       eaf_col = "AF", effect_allele_col = "REF", other_allele_col = "ALT", clump = TRUE)
dim(exposure_GSCAN_AI) # dim = (7,13)
# Get effects of instruments on outcome
outcome_Kunkle_AD = extract_outcome_data(snps = exposure_GSCAN_AI$SNP, outcomes = "ieu-b-2") # Kunkle,2019
# To detect if there is any IVs not available in outcome data
missing = exposure_GSCAN_AI %>% filter(!SNP %in% outcome_Kunkle_AD$SNP) # 0 SNP missing
# Harmonization of data
dat_AI = harmonise_data(exposure_dat = exposure_GSCAN_AI, outcome_dat = outcome_Kunkle_AD, action = 2)
# Perform MR
mr_AI_result = extract_MR(mr(dat_AI), "age_init", selected_method)
```

### 4) Smoking cessation
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
exposure_GSCAN_SC = read_exposure_data(paste0(work_data_path, "sumstats/sig_SmokingCessation.txt"), sep = "\t", 
                                       samplesize_col = "N", snp_col = "RSID", beta_col = "BETA", se_col = "SE", pval_col = "P", 
                                       eaf_col = "AF", effect_allele_col = "REF", other_allele_col = "ALT", clump = TRUE)
dim(exposure_GSCAN_SC) # dim = (8,13)
# Get effects of instruments on outcome
outcome_Kunkle_AD = extract_outcome_data(snps = exposure_GSCAN_SC$SNP, outcomes = "ieu-b-2") # Kunkle,2019
# To detect if there is any IVs not avSClable in outcome data
missing = exposure_GSCAN_SC %>% filter(!SNP %in% outcome_Kunkle_AD$SNP) # 1 SNP missing
# Find proxy SNP using LDlinkR
x = LDproxy(missing[1, 1], pop = "EUR", r2d = "r2", token = "6e1ccff03a31") 
eligible = x[x$R2 >= 0.8, ]
A = exposure_GSCAN_SC %>% filter(SNP %in% eligible$RS_Number) 
Common = outcome_Kunkle_AD %>% filter(SNP %in% A$SNP) # no common SNPs
# Harmonization of data
dat_SC = harmonise_data(exposure_dat = exposure_GSCAN_SC, outcome_dat = outcome_Kunkle_AD, action = 2)
# Perform MR
mr_SC_result = extract_MR(mr(dat_SC), "smk_cess", selected_method)
```

### Combine MR results
```{r}
mr_result = rbind(mr_SI_result, mr_CPD_result, mr_AI_result, mr_SC_result)
```

## 3. Assumption tests
### 1) F-statistics and R2
```{r}
# Calculate a per-SNP F statistic 
# Smoking initiation
N = 632802 ; k = dat_SI %>% filter(mr_keep == T) %>% nrow()
dat_SI_add = dat_SI %>% mutate(EAF2 = 1 - eaf.exposure, MAF = pmin(eaf.exposure, EAF2)) %>% 
  mutate(PVE = PVEfx(beta.exposure, MAF, se.exposure, N),
         FSTAT = ((N - 1 - 1)/1)*(PVE/(1 - PVE)) )
F_SI = ((N - 1)/k)*((sum(dat_SI_add$PVE))/(1 - sum(dat_SI_add$PVE))) 
# Cigarettes per day
N = 263954 ; k = dat_CPD %>% filter(mr_keep == T) %>% nrow()
dat_CPD_add = dat_CPD %>% mutate(EAF2 = 1 - eaf.exposure, MAF = pmin(eaf.exposure, EAF2)) %>% 
  mutate(PVE = PVEfx(beta.exposure, MAF, se.exposure, N),
         FSTAT = ((N - 1 - 1)/1)*(PVE/(1 - PVE)) )
F_CPD = ((N - 1)/k)*((sum(dat_CPD_add$PVE))/(1 - sum(dat_CPD_add$PVE))) 
# Age of initiation
N = 262990 ; k = dat_AI %>% filter(mr_keep == T) %>% nrow()
dat_AI_add = dat_AI %>% mutate(EAF2 = 1 - eaf.exposure, MAF = pmin(eaf.exposure, EAF2)) %>% 
  mutate(PVE = PVEfx(beta.exposure, MAF, se.exposure, N),
         FSTAT = ((N - 1 - 1)/1)*(PVE/(1 - PVE)) )
F_AI = ((N - 1)/k)*((sum(dat_AI_add$PVE))/(1 - sum(dat_AI_add$PVE))) 
# Smoking cessation
N = 312821 ; k = dat_SC %>% filter(mr_keep == T) %>% nrow()
dat_SC_add = dat_SC %>% mutate(EAF2 = 1 - eaf.exposure, MAF = pmin(eaf.exposure, EAF2)) %>% 
  mutate(PVE = PVEfx(beta.exposure, MAF, se.exposure, N),
         FSTAT = ((N - 1 - 1)/1)*(PVE/(1 - PVE)) )
F_SC = ((N - 1)/k)*((sum(dat_SC_add$PVE))/(1 - sum(dat_SC_add$PVE))) 
```

```{r message=FALSE, warning=FALSE}
mr_report(dat_SI, output_path = paste0(output_path, "TSMR_report/SmokingInitiation/"), study = "SI and AD")
mr_report(dat_CPD, output_path = paste0(output_path, "TSMR_report/CigarettesPerDay/"), study = "CPD and AD")
mr_report(dat_AI, output_path = paste0(output_path, "TSMR_report/AgeInitiation/"), study = "AI and AD")
mr_report(dat_SC, output_path = paste0(output_path, "TSMR_report/SmokingCessation/"), study = "SC and AD")
```


```{r}
# MR PRESSO to detect horizontal pleiotropy
presso = mr_presso(BetaOutcome = "beta.outcome", BetaExposure = "beta.exposure", SdOutcome = "se.outcome", 
                   SdExposure = "se.exposure", OUTLIERtest = TRUE, DISTORTIONtest = TRUE, data = dat_SI, 
                   NbDistribution = 1000, SignifThreshold = 0.05)
capture.output(print(presso), file = paste0(output_path, "MRPRESSO/SI_presso.txt"))

presso = mr_presso(BetaOutcome = "beta.outcome", BetaExposure = "beta.exposure", SdOutcome = "se.outcome", 
                   SdExposure = "se.exposure", OUTLIERtest = TRUE, DISTORTIONtest = TRUE, data = dat_CPD, 
                   NbDistribution = 1000, SignifThreshold = 0.05)
capture.output(print(presso), file = paste0(output_path, "MRPRESSO/CPD_presso.txt"))

presso = mr_presso(BetaOutcome = "beta.outcome", BetaExposure = "beta.exposure", SdOutcome = "se.outcome", 
                   SdExposure = "se.exposure", OUTLIERtest = TRUE, DISTORTIONtest = TRUE, data = dat_AI, 
                   NbDistribution = 1000, SignifThreshold = 0.05)
capture.output(print(presso), file = paste0(output_path, "MRPRESSO/AI_presso.txt"))

presso = mr_presso(BetaOutcome = "beta.outcome", BetaExposure = "beta.exposure", SdOutcome = "se.outcome", 
                   SdExposure = "se.exposure", OUTLIERtest = TRUE, DISTORTIONtest = TRUE, data = dat_SC, 
                   NbDistribution = 1000, SignifThreshold = 0.05)
capture.output(print(presso), file = paste0(output_path, "MRPRESSO/SC_presso.txt"))
```

# Part 2. Association check
```{r message=FALSE, warning=FALSE}
rm(list = ls())
lapply(paste('package:', names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE)
pacman::p_load(tidyverse, epiflow, writexl)
# Basic setups
work_data_path = '/Users/Mingzhou/Desktop/AD_Grant/smk_dem_MR/data/'
output_path = '/Users/Mingzhou/Desktop/AD_Grant/smk_dem_MR/output/'
# Source in useful functions
source("functions.R")
load(file = paste0(work_data_path, "modeling/final_europ_0124.rda"))
```

## 1. Smoking behaviors and cognition
```{r}
# Set up models
demo_basic = 'age + sex + education + last_cog_wave + PC1_5A + PC1_5B + PC1_5C + PC1_5D + PC1_5E'
health_behavior = paste0(demo_basic, ' + ever_drink + bmi + cesd + hypertension + diabetes + stroke')
gene_add = paste0(health_behavior, ' + APOE2010_bin')
```

### 1) Ordinal logistic regression check
```{r}
library(MASS)
# fit ordered logit model and store results 'm'
olm = polr(as.formula(paste('as.factor(cog_last_visit) ~ ever_smoke + ', demo_basic)), data = final_europ, Hess = TRUE)
summary(olm)
detach("package:MASS")
```

### 2) Back to multivariate regression
```{r}
cind_normal = final_europ %>% filter(cog_last_visit == 0 | cog_last_visit == 1) %>% 
  mutate(cind_bin = if_else(cog_last_visit == 0, 0, 1)) 
# dim = (7446,39)
dementia_normal = final_europ %>% filter(cog_last_visit == 0 | cog_last_visit == 2) %>% 
  mutate(dem_bin = if_else(cog_last_visit == 0, 0, 1))
# dim = (6637,39)
```

```{r message=FALSE, warning=FALSE}
smk_var_lst = c("ever_smoke", "current_smoke", "yrs_since_start", "age_stop_smk", "total_smk_yr")
lst_model = c('demo', 'health', 'apoe')

# CIND
for (i in 1:length(smk_var_lst)) {
  smk_var = smk_var_lst[i]
  f1 = as.formula(paste('cind_bin ~ ', smk_var, ' + ', demo_basic))
  demo = glm(f1, data = cind_normal, family = 'binomial')
  f2 = as.formula(paste('cind_bin ~ ', smk_var, ' + ', health_behavior))
  health = glm(f2, data = cind_normal, family = 'binomial')
  f3 = as.formula(paste('cind_bin ~ ', smk_var, ' + ', gene_add))
  apoe = glm(f3, data = cind_normal, family = 'binomial')
  or_result = make_OR_table(lst_model, 2, paste0('cind_', smk_var, '_'))
  if (i == 1) {cind_tbl = or_result} else {cind_tbl = rbind(cind_tbl, or_result)}
}
# Dementia
for (i in 1:length(smk_var_lst)) {
  smk_var = smk_var_lst[i]
  f1 = as.formula(paste('dem_bin ~ ', smk_var, ' + ', demo_basic))
  demo = glm(f1, data = dementia_normal, family = 'binomial')
  f2 = as.formula(paste('dem_bin ~ ', smk_var, ' + ', health_behavior))
  health = glm(f2, data = dementia_normal, family = 'binomial')
  f3 = as.formula(paste('dem_bin ~ ', smk_var, ' + ', gene_add))
  apoe = glm(f3, data = dementia_normal, family = 'binomial')
  or_result = make_OR_table(lst_model, 2, paste0('dem_', smk_var, '_'))
  if (i == 1) {dem_tbl = or_result} else {dem_tbl = rbind(dem_tbl, or_result)}
}
smk_cog_summary = rbind(cind_tbl, dem_tbl)
```

So here we decide to do smoking PGSs -> smoking -> CIND and AD PGS -> dementia -> smoking cessation

## 2. Smoking PGSs and smoking behaviors
```{r message=FALSE, warning=FALSE}
gene_adj_basic = 'age + sex + PC1_5A + PC1_5B + PC1_5C + PC1_5D + PC1_5E'
smk_var_bin = c("ever_smoke", "current_smoke")
smk_var_cont = c("age_start_smk", "yrs_since_start", "age_stop_smk", "total_smk_yr")
prs_lst = c("PRS_SI", "PRS_CPD", "PRS_AI", "PRS_SC")
# Binary
for (i in 1:length(smk_var_bin)) {
  smk_var = smk_var_bin[i]
  for (j in 1:length(prs_lst)) {
    prs_var = prs_lst[j]
    f1 = as.formula(paste(smk_var, ' ~ ', prs_var, ' + ', gene_adj_basic))
    prs_demo = glm(f1, data = final_europ, family = 'binomial')
    or_result = make_OR_table("prs_demo", 2, paste0(prs_var, '_', smk_var, '_'))
    if (i == 1 & j == 1) {prs_tbl_bin = or_result} else {prs_tbl_bin = rbind(prs_tbl_bin, or_result)}
  }
}
# Continuous
for (i in 1:length(smk_var_cont)) {
  smk_var = smk_var_cont[i]
  for (j in 1:length(prs_lst)) {
    prs_var = prs_lst[j]
    f1 = as.formula(paste(smk_var, ' ~ ', prs_var, ' + ', gene_adj_basic))
    prs_demo = lm(f1, data = final_europ)
    or_result = make_coeff_table_lm("prs_demo", 2, paste0(prs_var, '_', smk_var, '_'))
    if (i == 1 & j == 1) {prs_tbl_cont = or_result} else {prs_tbl_cont = rbind(prs_tbl_cont, or_result)}
  }
}
prs_tbl_cont$OR = NA
prs_tbl_cont$`95% CI_OR` = NA
pgs_smk_summary = rbind(prs_tbl_bin, prs_tbl_cont)
```

## 3. Smoking PGSs and cognition
```{r message=FALSE, warning=FALSE}
prs_lst = c("PRS_SI", "PRS_CPD", "PRS_AI", "PRS_SC")
lst_model = c('demo', 'health', 'apoe')

# CIND
for (i in 1:length(prs_lst)) {
  prs_var = prs_lst[i]
  f1 = as.formula(paste('cind_bin ~ ', prs_var, ' + ', demo_basic))
  demo = glm(f1, data = cind_normal, family = 'binomial')
  f2 = as.formula(paste('cind_bin ~ ', prs_var, ' + ', health_behavior))
  health = glm(f2, data = cind_normal, family = 'binomial')
  f3 = as.formula(paste('cind_bin ~ ', prs_var, ' + ', gene_add))
  apoe = glm(f3, data = cind_normal, family = 'binomial')
  or_result = make_OR_table(lst_model, 2, paste0('dem_', prs_var, '_'))
  if (i == 1) {cind_prs_tbl = or_result} else {cind_prs_tbl = rbind(cind_prs_tbl, or_result)}
}
# Dementia
for (i in 1:length(prs_lst)) {
  prs_var = prs_lst[i]
  f1 = as.formula(paste('dem_bin ~ ', prs_var, ' + ', demo_basic))
  demo = glm(f1, data = dementia_normal, family = 'binomial')
  f2 = as.formula(paste('dem_bin ~ ', prs_var, ' + ', health_behavior))
  health = glm(f2, data = dementia_normal, family = 'binomial')
  f3 = as.formula(paste('dem_bin ~ ', prs_var, ' + ', gene_add))
  apoe = glm(f3, data = dementia_normal, family = 'binomial')
  or_result = make_OR_table(lst_model, 2, paste0('dem_', prs_var, '_'))
  if (i == 1) {dem_prs_tbl = or_result} else {dem_prs_tbl = rbind(dem_prs_tbl, or_result)}
}
prs_cog_summary = rbind(cind_prs_tbl, dem_prs_tbl)
```

## Output results
```{r}
sheets_OR = list('smk_cog' = smk_cog_summary, 'pgs_smk' = pgs_smk_summary,
                 'pgs_cog' = prs_cog_summary)
write_xlsx(sheets_OR, path = paste0(output_path, 'association_0124.xlsx'))
```

# Part 3. One-sample MR
```{r message=FALSE, warning=FALSE}
rm(list = ls())
lapply(paste('package:', names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE)
pacman::p_load(tidyverse, epiflow, writexl, AER, modelsummary)
# Basic setups
work_data_path = '/Users/Mingzhou/Desktop/AD_Grant/smk_dem_MR/data/'
output_path = '/Users/Mingzhou/Desktop/AD_Grant/smk_dem_MR/output/'
# Source in useful functions
source("functions.R")
load(file = paste0(work_data_path, "modeling/final_europ_0124.rda"))
```

```{r}
# Separate by cognitive status
cind_normal = final_europ %>% filter(cog_last_visit == 0 | cog_last_visit == 1) %>% 
  mutate(cind_bin = if_else(cog_last_visit == 0, 0, 1)) 
# dim = (7446,39)
dementia_normal = final_europ %>% filter(cog_last_visit == 0 | cog_last_visit == 2) %>% 
  mutate(dem_bin = if_else(cog_last_visit == 0, 0, 1))
# dim = (6637,39)
gene_adj_basic = 'age + sex + PC1_5A + PC1_5B + PC1_5C + PC1_5D + PC1_5E + APOE2010_bin'
demo_basic = 'age + sex + education + last_cog_wave + PC1_5A + PC1_5B + PC1_5C + PC1_5D + PC1_5E'
health_behavior = paste0(demo_basic, ' + ever_drink + bmi + cesd + hypertension + diabetes + stroke')
gene_add = paste0(health_behavior, ' + APOE2010_bin')
```

## 1. CIND vs. Normal
```{r}
# Ever smoking
f0 = as.formula(paste0('cind_bin ~ ever_smoke', ' + ', gene_add))
f1 = as.formula(paste0('cind_bin ~ ever_smoke', ' + ', gene_adj_basic, ' | PRS_SI + ', gene_add))
# perform iv regression
glm_ever = glm(f0, data = cind_normal, family = "binomial")
iv_mod_ever = ivreg(f1, data = cind_normal)
m_list = list(GLM = glm_ever, IV = iv_mod_ever)
msummary(m_list)
summary(iv_mod_ever, diagnostics = TRUE)

# Current smoking
f0 = as.formula(paste0('cind_bin ~ current_smoke', ' + ', gene_add))
f2 = as.formula(paste0('cind_bin ~ current_smoke', ' + ', gene_adj_basic, ' | PRS_SC + ', gene_add))
glm_current = glm(f0, data = cind_normal, family = "binomial")
iv_mod_current = ivreg(f2, data = cind_normal)
m_list = list(GLM = glm_current, IV = iv_mod_current)
msummary(m_list)
summary(iv_mod_current, diagnostics = TRUE)

# Age stop smoking
f0 = as.formula(paste0('cind_bin ~ age_stop_smk', ' + ', gene_add))
f3 = as.formula(paste0('cind_bin ~ age_stop_smk', ' + ', gene_adj_basic, ' | PRS_SI + ', gene_add))
glm_agestop = glm(f0, data = cind_normal, family = "binomial")
iv_mod_agestop = ivreg(f3, data = cind_normal)
m_list = list(GLM = glm_agestop, IV = iv_mod_agestop)
msummary(m_list)
summary(iv_mod_agestop, diagnostics = TRUE)

# Total smoking years
f0 = as.formula(paste0('cind_bin ~ total_smk_yr', ' + ', gene_add))
f4 = as.formula(paste0('cind_bin ~ total_smk_yr', ' + ', gene_adj_basic, ' | PRS_SI + ', gene_add))
glm_smkyr = glm(f0, data = cind_normal, family = "binomial")
iv_mod_smkyr = ivreg(f4, data = cind_normal)
m_list = list(GLM = glm_smkyr, IV = iv_mod_smkyr)
msummary(m_list)
summary(iv_mod_smkyr, diagnostics = TRUE)
```



```{r}
# Ever smoking
f0 = as.formula(paste0('dem_bin ~ ever_smoke', ' + ', gene_add))
f1 = as.formula(paste0('dem_bin ~ ever_smoke', ' + ', gene_adj_basic, ' | PRS_SI + ', gene_add))
# perform iv regression
glm_ever = glm(f0, data = dementia_normal, family = "binomial")
iv_mod_ever = ivreg(f1, data = dementia_normal)
m_list = list(GLM = glm_ever, IV = iv_mod_ever)
msummary(m_list)
summary(iv_mod_ever, diagnostics = TRUE)

# Current smoking
f0 = as.formula(paste0('dem_bin ~ current_smoke', ' + ', gene_add))
f2 = as.formula(paste0('dem_bin ~ current_smoke', ' + ', gene_adj_basic, ' | PRS_SC + ', gene_add))
glm_current = glm(f0, data = dementia_normal, family = "binomial")
iv_mod_current = ivreg(f2, data = dementia_normal)
m_list = list(GLM = glm_current, IV = iv_mod_current)
msummary(m_list)
summary(iv_mod_current, diagnostics = TRUE)

# Age stop smoking
f0 = as.formula(paste0('dem_bin ~ age_stop_smk', ' + ', gene_add))
f3 = as.formula(paste0('dem_bin ~ age_stop_smk', ' + ', gene_adj_basic, ' | PRS_SI + ', gene_add))
glm_agestop = glm(f0, data = dementia_normal, family = "binomial")
iv_mod_agestop = ivreg(f3, data = dementia_normal)
m_list = list(GLM = glm_agestop, IV = iv_mod_agestop)
msummary(m_list)
summary(iv_mod_agestop, diagnostics = TRUE)

# Total smoking years
f0 = as.formula(paste0('dem_bin ~ total_smk_yr', ' + ', gene_add))
f4 = as.formula(paste0('dem_bin ~ total_smk_yr', ' + ', gene_adj_basic, ' | PRS_SI + ', gene_add))
glm_smkyr = glm(f0, data = dementia_normal, family = "binomial")
iv_mod_smkyr = ivreg(f4, data = dementia_normal)
m_list = list(GLM = glm_smkyr, IV = iv_mod_smkyr)
msummary(m_list)
summary(iv_mod_smkyr, diagnostics = TRUE)
```


